---
title: "Conditions"
format:
  html:
    embed-resources: true
    toc: true
execute:
  warning: false
editor_options: 
  chunk_output_type: console
editor: visual
---

We load the required packages for data manipulation and visualization.

```{r}
library(tidyverse)
library(ggthemes)
```

## Import

We continue to work with a subset of the "What a Waste" database.

We read the waste data from a CSV file.

```{r}
waste <- read_csv("data/processed/waste-city-level-sml.csv")
```

We will also use an example spreadsheet that was created by one of the course participants.

We read the faecal sludge solids data from an Excel file.

```{r}
solids <- readxl::read_excel("data/raw/TS_poo_2022.xlsx")
```

## Explore

We use `glimpse()` to view the structure and first few values of the waste dataset.

```{r}

```

We count the number of observations in each income category.

```{r}

```

## Transform

Conditional statements with mutate() & case_when() of dplyr R package.

### waste data

We calculate per capita waste generation and create descriptive income categories using `case_when()`. Then we define custom factor levels and convert both income variables to factors with the specified order.

```{r}


```

```{r}

```

### Faecal sludge solids data

We use `case_when()` to apply a correction factor of 100 to septic tank measurements, while keeping all other source types unchanged.

```{r}

```

## Visualize

### Categories as character

We create a boxplot with jittered points showing waste generation by income category, using character variables for the categories.

```{r}
#| eval: false
ggplot(data = waste_cat,
       mapping = aes(x = income_cat,
                     y = generation_kg_capita,
                     color = income_cat)) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(size = 3, width = 0.1, alpha = 0.3) +
    scale_color_colorblind() +
    labs(x = "Income group",
         y = "Waste generation (tons per capita per year)")
```

### Categories as factor

We create the same plot as above, but now using factor variables for the categories, which ensures the income groups appear in the correct economic order.

```{r}
#| eval: false
ggplot(data = waste_fct,
       mapping = aes(x = income_cat,
                     y = generation_kg_capita,
                     color = income_cat)) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(size = 3, width = 0.1, alpha = 0.3) +
    scale_color_colorblind() +
    labs(x = "Income group",
         y = "Waste generation (tons per capita per year)")
```

## Communicate

-   Let's render
