---
title: "Conditions"
format:
  html:
    embed-resources: true
    toc: true
execute:
  warning: false
editor_options: 
  chunk_output_type: console
editor: visual
---

```{r}
library(tidyverse)
library(readxl)
```

## Import

We are using another faecal sludge solids example dataset

```{r}
sludge <- read_xlsx("data/raw/faecal-sludge-analysis.xlsx")
```

## Task 1

1.  A mistake happened during data entry for sample id 16. Use `mutate()` and `case_when()` to change the ts value of 0.72 to 8.72.

```{r}

```

## Task 2

1.  Another mistake happened during data entry for sample id 6. Use `mutate()` and `case_when()` to change the system value of id 6 from "pit latrine" to "septic tank".

```{r}


```

## Task 3 (stretch goal)

1.  Add a new variable with the name `ts_cat` to the data frame that categorizes sludge samples into low, medium and high solids content.Use `mutate()` and `case_when()` to create the new variable.

-   samples with less than 15 g/L are categorized as low
-   samples with 15 g/L to 50 g/L are categorized as medium
-   samples with more than 50 g/L are categorized as high

```{r}


```
