---
title: "Dates"
format:
  html:
    embed-resources: true
    toc: true
execute:
  warning: false
editor_options: 
  chunk_output_type: console
editor: visual
---

We load the required packages for data manipulation and reading Excel files.

```{r}
library(tidyverse)
library(readxl)
```

## Dates

We read an Excel file containing dates in various formats.

```{r}

dates <- read_excel("data/raw/date-formats.xlsx")

```

In R and other programming languages, dates are stored as numbers. The number of days since 1970-01-01 is the ISO 8601 standard.

In Excel, dates are stored as numbers of days since 1900-01-01. In Excel, the date number 1 corresponds to "1900-01-01," but this system incorrectly considers 1900 as a leap year, which it is not. As a result, to correctly interpret date numbers that originate from systems like Excel, the origin "1899-12-30" is used to account for this discrepancy

We convert different date formats to proper date classes using functions from the lubridate package. We handle ISO format, US format (month/day/year), EU format (day/month/year), numeric dates from Excel, extract dates from datetime values, adjust time zones, and add today's date. The `OlsonNames()` function displays all available time zones.

```{r}

```

We demonstrate how dates are stored internally as numbers, showing the numeric value of today's date and converting the number 1 back to a date.

```{r}

```

We extract various components from a date: year, month (with French labels), quarter, week, day, day of week (with French labels), day of year, and week of year.

```{r}


```
