---
title: "Conditions"
format:
  html:
    embed-resources: true
    toc: true
execute:
  warning: false
editor_options: 
  chunk_output_type: console
editor: visual
---

```{r}
library(tidyverse)
library(ggthemes)
```

## Import

We continue to work with a subset of the "What a Waste" database.

```{r}
waste <- read_csv("data/processed/waste-city-level-sml.csv")
```

We will also use an example spreadsheet that was created by one of the course participants.

```{r}
solids <- readxl::read_excel("data/raw/TS_poo_2022.xlsx")
```

## Explore

```{r}
glimpse(waste)
```

```{r}
waste |> 
    count(income_id)
```

## Transform

Conditional statements with mutate() & case_when() of dplyr R package

### waste data

```{r}
#| eval: false

waste_cat <- waste |> 
    mutate(generation_kg_capita = generation_tons_year / population * 1000) |> 
    mutate(income_cat = case_when(
        income_id == "HIC" ~ "high income",
        income_id == "UMC" ~ "upper-middle income",
        income_id == "LMC" ~ "lower-middle income",
        income_id == "LIC" ~ "low income"
    ))


levels_income <- c("HIC", "UMC", "LMC", "LIC")

levels_income_cat <- c("high income", 
                       "upper-middle income",
                       "lower-middle income",
                       "low income")

waste_fct <- waste_cat |>
    mutate(income_id = factor(income_id, levels = levels_income)) |> 
    mutate(income_cat = factor(income_cat, levels = levels_income_cat)) |> 
    relocate(income_cat, .after = income_id)

```

```{r}
write_rds(x = waste_fct, file = "data/processed/waste-city-level-sml.rds")
```

### Faecal sludge solids data

```{r}
solids |> 
    mutate(total_solids_gL = case_when(
        source_type == "septic tank" ~ total_solids_gL * 100,
        .default = total_solids_gL
    ))
```

## Visualize

### Categories as character

```{r}
#| eval: true
ggplot(data = waste_cat,
       mapping = aes(x = income_cat, 
                     y = generation_kg_capita, 
                     color = income_cat)) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(size = 3, width = 0.1, alpha = 0.3) +
    scale_color_colorblind() +
    labs(x = "Income group",
         y = "Waste generation (tons per capita per year)") 
```

### Categories as factor

```{r}

#| eval: true
ggplot(data = waste_fct,
       mapping = aes(x = income_cat, 
                     y = generation_kg_capita, 
                     color = income_cat)) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(size = 3, width = 0.1, alpha = 0.3) +
    scale_color_colorblind() +
    labs(x = "Income group",
         y = "Waste generation (tons per capita per year)") 
```

## Communicate

-   Let's render
