---
title: "Dates"
format:
  html:
    embed-resources: true
    toc: true
execute:
  warning: false
editor_options: 
  chunk_output_type: console
editor: visual
---

```{r}
library(tidyverse)
library(readxl)
```

## Dates

```{r}

dates <- read_excel("data/raw/date-formats.xlsx")

```

In R and other programming languages, dates are stored as numbers. The number of days since 1970-01-01 is the ISO 8601 standard.

In Excel, dates are stored as numbers of days since 1900-01-01. In Excel, the date number 1 corresponds to "1900-01-01," but this system incorrectly considers 1900 as a leap year, which it is not. As a result, to correctly interpret date numbers that originate from systems like Excel, the origin "1899-12-30" is used to account for this discrepancy

```{r}
dates_class <- dates |> 
    mutate(date_iso = as_date(date_iso)) |> 
    mutate(date_us = mdy(date_us)) |> 
    mutate(date_eu = dmy(date_eu)) |> 
    mutate(date_num = as_date(date_num, origin = "1899-12-30")) |> 
    mutate(date = as_date(date_time)) |> 
    mutate(date_time_tz = with_tz(date_time, tzone = "Africa/Kampala")) |>
    mutate(today = today())

OlsonNames()
```

```{r}
as.numeric(today())
as_date(1)
```

```{r}
dates_class |> 
    select(today) |> 
    mutate(year = year(today)) |>
    mutate(month = month(today, label = TRUE, abbr = FALSE, locale = "fr_FR")) |> 
    mutate(quarter = quarter(today)) |>
    mutate(week = week(today)) |>
    mutate(day = day(today)) |>
    mutate(day_of_week = wday(today, label = TRUE, abbr = FALSE, locale = "fr_FR")) |>
    mutate(day_of_year = yday(today)) |>
    mutate(week_of_year = week(today)) 


```
