---
title: "Conditions"
format:
  html:
    embed-resources: true
    toc: true
execute:
  warning: false
editor_options: 
  chunk_output_type: console
editor: visual
---

```{r}
library(tidyverse)
library(ggthemes)
```

## Import

We continue to work with a subset of the "What a Waste" database.

```{r}
waste <- read_csv("data/processed/waste-city-level-sml.csv")
```

We will also use an example spreadsheet that was created by one of the course participants.

```{r}
solids <- readxl::read_excel("data/raw/TS_poo_2022.xlsx")
```

## Explore

```{r}

```

```{r}

```

## Transform

Conditional statements with mutate() & case_when() of dplyr R package

### waste data

```{r}


```

```{r}

```

### Faecal sludge solids data

```{r}

```

## Visualize

### Categories as character

```{r}
#| eval: false
ggplot(data = waste_cat,
       mapping = aes(x = income_cat, 
                     y = generation_kg_capita, 
                     color = income_cat)) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(size = 3, width = 0.1, alpha = 0.3) +
    scale_color_colorblind() +
    labs(x = "Income group",
         y = "Waste generation (tons per capita per year)") 
```

### Categories as factor

```{r}
#| eval: false
ggplot(data = waste_fct,
       mapping = aes(x = income_cat, 
                     y = generation_kg_capita, 
                     color = income_cat)) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(size = 3, width = 0.1, alpha = 0.3) +
    scale_color_colorblind() +
    labs(x = "Income group",
         y = "Waste generation (tons per capita per year)") 
```

## Communicate

-   Let's render
